Optimal (Multiway) Spatial Joins

### 2024-05-25 msj.tex
yufei

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ (defined in \eqref{eqn:contained}), where the rectangles are sorted by their right boundaries. Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$. Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

@@

We are particularly concerned with the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$. The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties for these sets.

@@

@@

>>>>

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries. Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$. Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

@@

Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$. The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties of these sets.

@@

\extraspacing {\bf Set $\bm{H^*}$.} Fix any $i \in [k-2]$. For each segment $h' \in H'$, define
\myeqn{
    \minleft(h') &=& \min_{v \in V: h' \cap v \ne \emptyset} \text{$x$-coordinate of $v$}. 
}
We set $\minleft(h')$ to $\infty$, if $h'$ intersects no segment in $V$. When $\minleft(h') \ne \infty$, assuming $h' = [x_1, x_2] \times y$, we introduce a horizontal segment
\myeqn{
    h^* &=& [x_1, \minleft(h')] \times y. \nn
}
and call it the {\em minimal segment} of $r'$.

Next, we construct from $H'$ a new set of rectangles: 
\myeqn{
    H^* = \set{h^* \mid h' \in H' \text{ and its minimal segment $h^*$ exists}}. \label{eqn:hv:type2:H*}
}
Computing $H^*$ from $H'$ and $V$ is an instance of Problem $\mathscr{B}$, and can be achieved in $O(n \log n)$ as per the discussion from Section~\ref{sec:bricks}.

We will utilize the sets $\contained_{R_{k-2}}(h^*)$ of the segments $h^*$ in $H^*$, where $\contained_{R_{k-2}}(h^*)$ --- defined in \eqref{eqn:contained} --- is the set of rectangles in $R_{k-2}$ containing $h^*$. The next lemma is proven in Appendix~\ref{app:hv}.

\begin{lemma} \label{lmm:hv:type2:contained-h*}
    Both statements below are true: 
    \myenums{
        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.

@@

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}
